---
title: "Reproducible Research: Peer Assessment 1"
output: 
  html_document:
    keep_md: true
---

```{r,echo=FALSE,results='hide',warning=FALSE,message=FALSE}
library(dplyr)
library(ggplot2)
library(knitr)
```

## Loading and processing the data

**1-Load the data**
```{r,echo=TRUE,results='markup',message=TRUE}
if(!file.exists("C:/Users/Super/Desktop/RepData_PeerAssessment1/activity.csv")){
  unzip("C:/Users/Super/Desktop/RepData_PeerAssessment1/activity.zip")
}
activityData <- read.csv("C:/Users/Super/Desktop/RepData_PeerAssessment1/activity.csv",header = TRUE)
head(activityData,4)
```
**2.Process the data**
```{r}
#activityData$interval <- strptime(gsub("([0-9]{1,2})([0-9]{2})", "\\1:\\2", activityData$interval), format='%H:%M')
```

## What is mean total number of steps taken per day?

**1.Calculating total number of steps taken per day**
```{r,echo=TRUE}
totalSteps <- activityData %>% select(date,steps) %>% group_by(date) %>%
  summarize(Total.Steps = sum(steps,na.rm = TRUE))
head(totalSteps,3)
```
**2.Make a histogram of the total number of steps taken each day**
```{r histogram_total_steps, echo=TRUE}
hist(totalSteps$Total.Steps,main = "Total Number Of Steps Each Day",xlab = "Steps",ylab = "Frequency",col = "orange")
```
**3.Calculating mean and median of total number of steps taken daily**
```{r,echo=TRUE}
stepsMean <- mean(totalSteps$Total.Steps)
stepsMedian <- median(totalSteps$Total.Steps)
```
* Mean of total steps taken daily = `r stepsMean`
* Median of total steps taken daily = `r stepsMedian`

## What is the average daily activity pattern?

**1.Make a time series plot of the 5-minute interval and the average number of steps taken daily**
```{r,echo=TRUE}
averageOfActivity <- activityData %>% select(interval,steps) %>% na.omit() %>%
  group_by(interval) %>% summarize(Average.Steps = mean(steps))
head(averageOfActivity,3)
```

```{r average_steps_daily, echo=TRUE}
plot(x=averageOfActivity$interval,y=averageOfActivity$Average.Steps,type="l",
main = "Average Number Of Steps Taken Daily",xlab = "Interval",ylab = "Average Of Steps",col="purple")
```

**2.Which 5-minute interval contains maximum number of steps?**
```{r,echo=TRUE}
maxSteps <- averageOfActivity[which(averageOfActivity$Average.Steps==max(averageOfActivity$Average.Steps)),]
maxSteps
```
* Maximum steps at interval **835** which is **206.1698 steps**

## Imputting Missing Values

**1.Calculating total number of missing values in the dataset**
```{r,echo=TRUE}
totalMissingValues <- sum(is.na(activityData$steps))
```
* Total number of missing values `r totalMissingValues`

**2.Devise a strategy for filling in all of the missing values in the dataset.**

* Iam using **mean** function to replace all the missing values in the dataset.
```{r,echo=TRUE}
replacingMissingValues <- function(x) replace(x,is.na(x),mean(x,na.rm = TRUE))
```

**3.Create a new dataset that is equal to the orignal dataset but with the missing data filled in**
```{r,echo=TRUE}
newDataset <- activityData %>% group_by(interval) %>% mutate(steps=replacingMissingValues(steps))
head(newDataset,3)
```
**4(i).Make a histogram of the total number of steps taken each day**
```{r new_total_steps_taken_daily, echo=TRUE}
totalStepsDaily <- newDataset %>% group_by(date) %>% summarize(Total.Steps = sum(steps))

hist(totalStepsDaily$Total.Steps,main = "Total Number Of Steps Taken Daily",xlab = "Total Steps",ylab = "Frequency",col = "wheat")
```
**4(ii).Calculating mean and median total number of steps taken each day**
```{r,echo=TRUE}
newMean <- mean(totalStepsDaily$Total.Steps)
newMedian <- median(totalStepsDaily$Total.Steps)
```
* **Mean** = `r newMean` and **median** is = `r newMedian`

**4(iii).Do these values differ from the estimates from the first part of the dataset**
```{r,echo=TRUE}
oldMean <- mean(totalSteps$Total.Steps)
oldMedian <- median(totalSteps$Total.Steps)
```
* Yes values of mean and median are different from the first part of the assignment
  * old mean = `r oldMean` and new mean = `r newMean`
  * old median = `r oldMedian` and new median = `r newMedian`

## Are there differences in activity patterns between weekdays and weekends?

**1.Create a new factor variable in the dataset with two levels – “weekday” and “weekend” indicating whether a given date is a weekday or weekend day**
```{r,echo=TRUE}
newDataset$date <- as.Date(newDataset$date)
newDataset$weekday <- weekdays(newDataset$date)
newDataset$weekend <- ifelse(newDataset$weekday=="Saturday" | newDataset$weekday=="Sunday","weekend","weekday")
```

**2.Make a panel plot containing a time series plot of the 5-minute interval (x-axis) and the average number of steps taken, averaged across all weekday days or weekend days (y-axis)**
```{r panel_plot, echo=TRUE}
dataWendWday <- aggregate(newDataset$steps,by=list(newDataset$weekend,newDataset$interval),na.omit(mean))

names(dataWendWday) <- c("weekend","interval","steps")

ggplot(dataWendWday,aes(x=interval,y=steps,color=weekend))+geom_line()+facet_grid(weekend~.)+xlab("Interval")+ylab("Average Number Of Steps")+ggtitle("Comparison Of Average Number Of Steps In Each Interval")
```